An efficient acyclic contact planner for multiped robots

### 2017-01-12 body/contact.tex
updated figure on julien s advice, integrated [CONTACT] modifs by Andrea

\subsection{\repadp{Contact}{Definition of a contact} generator}
\label{sec:single_contact}

@@

\repadp{The contact generation treats each limb $R^k$ independently.}{A contact generator is defined for a given limb $R^k$, and only affects the DoFs of $R^k$.}
Given a \repadp{configuration of the root and the other limbs}{robot configuration}, it computes a configuration $\mathbf{q}^k$ such that $R^k$ is in contact and the robot $R$ is in static equilibrium:
\begin{equation}
\label{eq:contact_gen}
	\mathbf{q}^{\overline{k}}  \longrightarrow \mathbf{q}^k, (\mathbf{q}^{k} \oplus \mathbf{q}^{\overline{k}}) \in  C_{Equil} \wedge \mathbf{q}^k \in  C_{Contact}^k 
\end{equation}

Contact generation is typically adressed by randomly sampling a limb configuration, before projecting the \deladp{considered} effector onto the closest surface with an inverse kinematics solver.
This process is repeated until either a solution for Eq.~\ref{eq:contact_gen} is found, or the generator fails, according to a user-defined maximum number of trials.
We favor a modified implementation of this naive approach, more computationally efficient, introduced in our previous work~\citep{Tonneau2014}.

@@

%~ This approach has two main advantages.
%~ First, this allows us to select a large number of candidates in a single proximity request.
%~ Having several candidates is interesting, because it allows to compare them using a user-selected heuristic $h$, thus obtaining
%~ a locally-optimal candidate.
%~ Furthermore, the fact that the candidates are already close to contact increases the odds that the inverse kinematics will converge to a valid solution in a small number of iterations.
%~ Regarding convergence, it is immediate to verify that as $N$ grows, the probability of finding a solution if it exists converges to 1.
%~ $N$ is a parameter that allows us to specify the trade-off between exhaustiveness and efficiency.
The reader is invited to refer to our previous work for an extensive discussion on the benefits of this approach, and the optimal choice 
of the \repadp{parameter $N$}{$N$ parameter}~\citep{Tonneau2014}. Our criterion to assert efficiently the static equilibrium of the robot, as well as the heuristics $h$ \deladp{used} are detailed in Appendix~\ref{sec:heuristics}.

@@

First, the root path $\mathbf{q}^0(t)$ is discretized into a sequence of $j$ key configurations:  
\begin{equation*}
	\mathbf{Q}^0 = [\mathbf{q}^0_{0}; \mathbf{q}^0_{i}; ..., \mathbf{q}^0_{j-1}]
\end{equation*} 
where $\mathbf{q}^0_{0}$ and $\mathbf{q}^0_{j-1}$ are the start and goal configurations. %To ensure continuity in the contact transition phases,
Each root configuration of $\mathbf{Q}^0$ is then extended into a whole-body configuration such that:
\begin{itemize} 
\item At most one contact is not maintained (\textit{broken}) between two consecutive configurations.
\item \addadp{At most one contact is added between two consecutive configurations.} 
\item Each configuration is in static equilibrium.
\item Each configuration is collision-free.
\end{itemize}

@@

%~ Figure~\ref{fig:break_contact} illustrates the contact-persistence strategy.
If kinematically possible, a limb in contact at step $i-1$ remains in contact at step $i$ (Figure~\ref{fig:break_contact}). 
%~ The contact is broken if an inverse-kinematics solver fails to find a collision-free limb configuration that satisfies joint limits. 
%~ The solver is directly provided by the HPP software.
Otherwise the contact is broken and a collision-free configuration is assigned to the limb.
If two or more contacts \repadp{cannot be maintained between two consecutive configurations}{are broken in a single step}, one or more intermediate \repadp{configurations}{steps} are added.
%~ For these steps the root configuration is the same as for the previous step, with the difference that
%~ one faulty contact is repositioned, in the hope that it will not be broken at the next step.

\begin{figure}[t]
\centering
  \begin{overpic}[width=0.9\linewidth]{figures/break_contact}
		\put (0,4) {1} 
		\put (25,4) {2} 
		\put (50,4) {3} 
		\put (76,4) {4} 
		%~ \put (68,58) {3.a)} 
		%~ \put (5,27) {3.b)} 
		%~ \put (37,27) {4.a)} 
		%~ \put (68,27) {4.b)} 
	\end{overpic}
\caption{Contacts are maintained if joint limits and collisions constraints are respected (2). They are broken otherwise(3,4).}
		   \label{fig:break_contact}
\end{figure}

@@

%\subsection{Generation of a contact configuration}  
\subsubsection{Creating contacts:}
Contacts are created using \deladp{three rules:} \addadp{a FIFO approach: we try first to create a contact with the limb that has been contact-free the longest. If the contact creation was not successful for a limb, the limb is pushed on top of the queue, and will only be tried again after the others.}
\begin{enumerate}
\item \deladp{At most one contact creation happens between two consecutive steps. }
%~ This increases the odds that the interpolation between the two steps be feasible;
%\adnote{Before you said that you ensured that transition between two states is feasible...}
\item \deladp{A contact is validated if and only if the resulting configuration is in static equilibrium;}
\item \deladp{We use a FIFO approach:  we always try first to create a contact with the limb that has been contact-free the longest. If the contact creation
was not successful for a limb, the limb is pushed on top of the queue, and will only be tried again after the others.}
\end{enumerate}

\repadp{This set of rules}{The set of rules defined in this Section} defines a deterministic contact planning algorithm.
The determinism of this algorithm makes it efficient, at the cost of not being probabilistically complete.
\deladp{However, in practice the planner is successful in the large majority of cases, as discussed in Section~\ref{sec:perf}.}

>>>>

\subsection{Contact generator}
\label{sec:single_contact}

@@

The contact generation treats each limb $R^k$ independently.
Given a configuration of the root and the other limbs, it computes a configuration $\mathbf{q}^k$ such that $R^k$ is in contact and the robot $R$ is in static equilibrium:
\begin{equation}
\label{eq:contact_gen}
	\mathbf{q}^{\overline{k}}  \longrightarrow \mathbf{q}^k, (\mathbf{q}^{k} \oplus \mathbf{q}^{\overline{k}}) \in  C_{Equil} \wedge \mathbf{q}^k \in  C_{Contact}^k 
\end{equation}

Contact generation is typically adressed by randomly sampling a limb configuration, before projecting the effector onto the closest surface with an inverse kinematics solver.
This process is repeated until either a solution for Eq.~\ref{eq:contact_gen} is found, or the generator fails, according to a user-defined maximum number of trials.
We favor a modified implementation of this naive approach, more computationally efficient, introduced in our previous work~\citep{Tonneau2014}.

@@

%~ This approach has two main advantages.
%~ First, this allows us to select a large number of candidates in a single proximity request.
%~ Having several candidates is interesting, because it allows to compare them using a user-selected heuristic $h$, thus obtaining
%~ a locally-optimal candidate.
%~ Furthermore, the fact that the candidates are already close to contact increases the odds that the inverse kinematics will converge to a valid solution in a small number of iterations.
%~ Regarding convergence, it is immediate to verify that as $N$ grows, the probability of finding a solution if it exists converges to 1.
%~ $N$ is a parameter that allows us to specify the trade-off between exhaustiveness and efficiency.
The reader is invited to refer to our previous work for an extensive discussion on the benefits of this approach, and the optimal choice 
of the parameter $N$~\citep{Tonneau2014}. Our criterion to assert efficiently the static equilibrium of the robot, as well as the heuristics $h$ are detailed in Appendix~\ref{sec:heuristics}.

@@

First, the root path $\mathbf{q}^0(t)$ is discretized into a sequence of $j$ key configurations:  
\begin{equation*}
	\mathbf{Q}^0 = [\mathbf{q}^0_{0}; \mathbf{q}^0_{i}; ..., \mathbf{q}^0_{j-1}]
\end{equation*} 
where $\mathbf{q}^0_{0}$ and $\mathbf{q}^0_{j-1}$ are the start and goal configurations. %To ensure continuity in the contact transition phases,
Each root configuration of $\mathbf{Q}^0$ is then extended into a whole-body configuration such that:
\begin{itemize} 
\item At most one contact is not maintained (\textit{broken}) between two consecutive configurations.
\item At most one contact is added between two consecutive configurations.
\item Each configuration is in static equilibrium.
\item Each configuration is collision-free.
\end{itemize}

@@

%~ Figure~\ref{fig:break_contact} illustrates the contact-persistence strategy.
If kinematically possible, a limb in contact at step $i-1$ remains in contact at step $i$ (Figure~\ref{fig:break_contact}). 
%~ The contact is broken if an inverse-kinematics solver fails to find a collision-free limb configuration that satisfies joint limits. 
%~ The solver is directly provided by the HPP software.
Otherwise the contact is broken and a collision-free configuration is assigned to the limb.
If two or more contacts can't be maintained between two consecutive configurations, one or more intermediate configurations are added.
%~ For these steps the root configuration is the same as for the previous step, with the difference that
%~ one faulty contact is repositioned, in the hope that it will not be broken at the next step.

\begin{figure}[t]
\centering
  \begin{overpic}[width=0.9\linewidth]{figures/break_contact}
		\put (0,4) {1} 
		\put (25,4) {2} 
		\put (50,4) {3} 
		\put (76,4) {4} 
		%~ \put (68,58) {3.a)} 
		%~ \put (5,27) {3.b)} 
		%~ \put (37,27) {4.a)} 
		%~ \put (68,27) {4.b)} 
	\end{overpic}
\caption{Contacts are maintained if joint limits and collisions constraints are respected (2). They are broken otherwise(3,4). The green line represents the root path.}
		   \label{fig:break_contact}
\end{figure}

@@

%\subsection{Generation of a contact configuration}  
\subsubsection{Creating contacts:}
Contacts are created using a FIFO approach: we try first to create a contact with the limb that has been contact-free the longest. If the contact creation was not successful, the limb is pushed on top of the queue, and will only be tried again after the others.
%~ \begin{enumerate}
%~ \item \deladp{At most one contact creation happens between two consecutive steps. }
%~ \item \deladp{A contact is validated if and only if the resulting configuration is in static equilibrium;}
%~ \item \deladp{We use a FIFO approach:  we always try first to create a contact with the limb that has been contact-free the longest. If the contact creation
%~ was not successful for a limb, the limb is pushed on top of the queue, and will only be tried again after the others.}
%~ \end{enumerate}

This set of rules defines a deterministic contact planning algorithm.
The determinism of this algorithm makes it efficient, at the cost of not being probabilistically complete.
%~ \deladp{However, in practice the planner is successful in the large majority of cases, as discussed in Section~\ref{sec:perf}.}

### 2017-01-12 body/intro.tex
[Contact planning] + cosmetics in overview intro and rbprm
updated figure on julien s advice, integrated [CONTACT] modifs by Andrea

% !TEX root =  ../main.tex

@@

@@

\newcommand{\Pa}{$\mathcal{P}_1$ }
\newcommand{\Pb}{$\mathcal{P}_2$ }

@@

\begin{figure*}
  \centering
  \begin{overpic}[width=0.8\linewidth]{figures/workflow}
    %~ \put (6.4,1.8) {\normalsize{Request}} 
    \put (30,10.8) {\large{\color{white}\Pa} }
    %~ \put (31.6,2) {\scriptsize{\color{white}RB-RRT}} 
    \put (66.4,10.8) {\large{\color{white}\Pb} }
    %~ \put (72.5,2) {\tiny{\color{white}Contact}} 
  \end{overpic}
  \vspace{-1em}
  \caption{
    Overview of our two-stage framework. Given a path request between start and goal positions (left image), \Pa is the problem of computing a guide path in the space
    of \textit{equilibrium feasible} root configurations. We achieve this by defining a geometric condition, the \textit{reachability condition} (abstracted with the transparent cylinders on the middle image). \Pb is then the problem of extending the path into a discrete sequence of contact configurations using an iterative algorithm (right image).}
  \label{fig:framework}
\end{figure*}

@@

Finally, to further demonstrate the validity of our approach, we show that the generated contact plans  can be successfully executed (problem  $\mathcal{P}_3$), either in simulation or on the real HRP-2 robot. For HRP-2, we detail the complete computation times to address sequentially the three problems, and compare them to related work, demonstrating that our method is order of magnitude faster.

>>>>

% !TEX root =  ../main.tex
\newcommand{\Pa}{$\mathcal{P}_1$ }
\newcommand{\Pb}{$\mathcal{P}_2$ }

@@

\begin{figure*}
  \centering
  \begin{overpic}[width=0.8\linewidth]{figures/workflow}
    %~ \put (6.4,1.8) {\normalsize{Request}} 
    \put (30,10.8) {\large{\color{white}\Pa} }
    %~ \put (31.6,2) {\scriptsize{\color{white}RB-RRT}} 
    \put (66.4,10.8) {\large{\color{white}\Pb} }
    %~ \put (72.5,2) {\tiny{\color{white}Contact}} 
	\put (15,1) {a} 
	\put (50,1) {b} 
	\put (85,1) {c} 
  \end{overpic}
  \vspace{-1em}
  \caption{
    Overview of our two-stage framework. Given a path request between start and goal positions (left image), \Pa is the problem of computing a guide path in the space
    of \textit{equilibrium feasible} root configurations. We achieve this by defining a geometric condition, the \textit{reachability condition} (abstracted with the transparent cylinders on the middle image). \Pb is then the problem of extending the path into a discrete sequence of contact configurations using an iterative algorithm (right image).}
  \label{fig:framework}
\end{figure*}

@@

@@

@@

To further demonstrate the validity of our approach, we show that the generated contact plans  can be successfully executed (problem  $\mathcal{P}_3$), either in simulation or on the real HRP-2 robot. For HRP-2, we detail the complete computation times to address sequentially the three problems, and compare them to related work, demonstrating that our method is orders of magnitude faster.

Finally, we provide an extensive discussion on the consequences of our approach in terms of efficiency and completeness regarding the contact planning problem.

### 2017-01-12 body/overview.tex
updated figure on julien s advice, integrated [CONTACT] modifs by Andrea

Figure~\ref{fig:framework} illustrates our workflow.
\Pa and \Pb are addressed sequentially: we first plan a root guide path, before
extending it into a sequence of configurations in static equilibrium.

@@

\medskip
$C_{Contact}$ is the set of whole body-configurations in contact and collision-free.
$C_{Contact}^k \subset C_{Contact}$ is the set of whole body-configurations where at least $R^k$ is in contact.

>>>>

Figure~\ref{fig:framework} illustrates our workflow.
\Pa and \Pb are addressed sequentially: From a given problem (a) we first plan a root guide path (b), before
extending it into a sequence of static equilibrium configurations(c). In the case where step c fails,
our framework invalidates the computed guide path, and restarts the planning from b.

@@

\medskip
Finally we define some relevant of the configuration space $C$.
$C_{Contact}$ is the set of whole body-configurations in contact and collision-free.
$C_{Contact}^k \subset C_{Contact}$ is the set of whole body-configurations where at least $R^k$ is in contact.

### 2017-01-12 body/rbprm.tex
updated figure on julien s advice, integrated [CONTACT] modifs by Andrea

\textbf{sufficient condition\addadp{s}:}
Similarly we can define an inner approximation \mbox{$\mathcal{C}^0_{\textrm{\it Suf}} \subset $ $C_{Contact}^0$} by considering a bounding volume $B^{\textrm{\it Suf}}$ encompassing the whole robot in a given pose, except for the effector surfaces in contact. 
\begin{equation}
\mathcal{C}^0_{\textrm{\it Suf}} = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \wedge B^{\textrm{\it Suf}}(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}

>>>>

\textbf{sufficient condition:}
Similarly we can define an inner approximation \mbox{$\mathcal{C}^0_{\textrm{\it Suf}} \subset $ $C_{Contact}^0$} by considering a bounding volume $B^{\textrm{\it Suf}}$ encompassing the whole robot in a given pose, except for the effector surfaces in contact. 
\begin{equation}
\mathcal{C}^0_{\textrm{\it Suf}} = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \wedge B^{\textrm{\it Suf}}(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
